P3064R2: How to Avoid OOTA Without Really Trying

### 2024-09-01 paper/ootavacuous.tex
paper/ootavacuous: Summarize "What Needs to Change" section
Signed-off-by: Paul E. McKenney <paulmck@kernel.org>

This paper shows that compiler-based C++ implementations subject to
reasonable constraints on how they treat accesses to atomic objects
cannot exhibit out-of-thin-air (OOTA) cycles.
It follows that these implementations need to take almost no special
actions to avoid OOTA.
In fact, for many compilers the constraints boil down to a simple
``Don't invent or duplicate atomic accesses'', which the compiler
probably wouldn't do anyway.

We begin with a summary of what needs to change and a brief overview of
the OOTA problem, followed by an equally brief summary of prior work in
this area, and ending with a quick overview of the \co{herd7} tool that
will be used to evaluate litmus tests.

\subsection{What Needs to Change?}
\label{sec:What Needs to Change?}

What changes are needed to prevent OOTA from C++ implementations using
traditional compilers and traditional hardware, including CPUs and
GPGPUs?

\begin{itemize}
\item	To the standard:
	No changes.
	The indefinite prohibition of out-of-thin-air values needs
	to remain.
\item	To implementations:
	No changes.
	Future implementations must avoid optimizations that invent,
	duplicate, or repurpose atomic loads.
	However, such optimizations must be avoided for other reasons,
	including preserving basic laws of arithmetic
	(see Appendix~\ref{app:Inventing Atomic Loads} for more detail).
\item	To user code:
	No changes.
	Users must avoid undefined behavior in portable code, just as
	is the case now.
\end{itemize}

In short, no changes are needed because current implementations already
avoid OOTA cycles.
The point of this paper is to give a semi-formal demonstration of
why this is the case.

>>>>

This paper shows that compiler-based C++ implementations subject to
reasonable constraints on how they treat accesses to atomic objects
cannot exhibit out-of-thin-air (OOTA) cycles.
It follows that these implementations need to take almost no special
actions to avoid OOTA.
In fact, for many compilers the constraints boil down to a simple
``Don't invent or duplicate atomic accesses'', which the compiler
probably wouldn't do anyway.
Absolutely no changes are required to either the standard or to user code.

We begin with a brief overview of the OOTA problem followed by an equally
brief summary of prior work in this area.
